---
title: "Assignment 4 Report"
author: "An Bui"
date: "11/10/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r set up libraries and data, include = FALSE}
# load packages
library(tidyverse)
library(vcdExtra)
library(kableExtra)

# read in .csv and get in tidy format
abund_master <- read_csv("lobster_size_abundance.csv")
abund <- as.data.frame(abund_master) %>% 
  expand.dft(freq = "COUNT")

traps_master <- read_csv("lobster_traps.csv")
traps <- as.data.frame(traps_master) %>% 
  expand.dft(freq = "TRAPS")
```

#### Ideas:  
1. box and whisker for abundance and fishing pressure (because sample size is so big that median is probably a better indicator of central value than mean)? or jitter?  
2. table of mean lobster sizes (or box and whisker? code below)
3. At Isla Vista and Naples Reef, the two protected MPA sites (with zero fishing pressure), how do lobster sizes
in 2012 and 2017 compare? At the non-MPA sites? -> t-test for IVEE and NAPL, ANOVA for AQUE, MOHK, CARP
4. What proportion of observed lobsters at each site are above the legal minimum? Does that proportion differ significantly across the 5 sites? -> chi square  

Thoughts are that we can just make a bunch of stuff and go through and pick what's useful and what's not?

##### exploratory things
```{r summary table, include = FALSE}
abund_table <- abund %>% 
  group_by(SITE) %>% 
  summarize(
    mean_size = round(mean(SIZE), 2),
    median_size = round(median(SIZE), 2),
    count = length(SIZE)
  ) %>% 
  mutate(site_full = case_when(
      SITE == "AQUE" ~ "Arroyo Quemado",
      SITE == "CARP" ~ "Carpinteria",
      SITE == "IVEE" ~ "Isla Vista",
      SITE == "MOHK" ~ "Mohawk Reef",
      SITE == "NAPL" ~ "Naples Reef"
    )
  ) %>% 
  select(site_full, mean_size, median_size, count)

abund_table_columns = c("Site", "Mean size", "Median size", "Sample size")

kable(abund_table,
      col.names = abund_table_columns)

trap_table <- traps %>% 
  group_by(SITE) %>% 
  filter(SITE == "AQUE" | SITE == "MOHK" | SITE == "CARP") %>% 
  summarize(
    count = length(SITE)
  )

trap_table
```

##### code for random shit
```{r abundance histogram, include = FALSE}
# just wanted to see what this looked like
abund_hist <- ggplot(abund, aes(x = SIZE)) +
  geom_histogram() +
  facet_wrap(~ SITE)

abund_hist
```

```{r box and whisker of size by site, include = FALSE}
abund_size_box <- ggplot(abund, aes(x = SITE, y = SIZE)) +
  geom_boxplot()

abund_size_box
```

```{r box and whisker of size by site in 2017 only, include = FALSE}
abund_2017 <- abund %>% 
  filter(YEAR == "2017")

size_box_2017 <- ggplot(abund_2017, aes(x = SITE, y = SIZE)) +
  geom_boxplot()

size_box_2017
```

```{r table of size by site in 2017 only, include = FALSE}
size_table_2017 <- abund_2017 %>% 
  group_by(SITE) %>% 
  summarize(
    mean = round(mean(SIZE), 2),
    median = round(median(SIZE), 2),
    sd = round(sd(SIZE), 2),
    n = length(SIZE)
    ) %>% 
  mutate(
    full_site = case_when(
      SITE == "AQUE" ~ "Arroyo Quemado",
      SITE == "IVEE" ~ "Isla Vista", 
      SITE == "MOHK" ~ "Mohawk",
      SITE == "CARP" ~ "Carpinteria",
      SITE == "NAPL" ~ "Naples Reef"
    )
  ) %>% 
  select(full_site, mean, median, sd, n)

size_table_2017_columns = c("Site", "Mean length", "Median length", "Standard deviation", "Sample size")

kable(size_table_2017,
      col.names = size_table_2017_columns) %>% 
  kable_styling(bootstrap_options = "striped")
```

##### stats n shit: IVEE and NAPL (part 3)
```{r summary stats, include = FALSE}
# Are mean lobster sizes significantly different in 2012 and 2017 for IVEE? For NAPL?
# H0: means are not significantly different in 2012 and 2017 for IVEE/NAPL.
# HA: means are significantly different in 2012 and 2017 for IVEE/NAPL.

# 1. establish normality: is 2012 IVEE/2017 IVEE and 2012 NAPL/2017 NAPL data normal?
# 2. test for equal variances
# 3. run a t-test

abund_ivee <- abund %>% 
  filter(YEAR == "2012" | YEAR == "2017",
         SITE == "IVEE")

abund_napl <- abund %>% 
  filter(YEAR == "2012" | YEAR == "2017",
         SITE == "NAPL")

abund_both <- abund %>% 
  filter(YEAR == "2012" | YEAR == "2017",
         SITE == "IVEE" | SITE == "NAPL")

abund_ivee_table <- abund_ivee %>% 
  group_by(YEAR) %>% 
  summarize(
    mean_size = mean(SIZE),
    median_size = median(SIZE),
    sd_size = sd(SIZE),
    sample_size = length(SIZE)
  )
abund_ivee_table

abund_napl_table <- abund_napl %>% 
  group_by(YEAR) %>% 
  summarize(
    mean_size = mean(SIZE),
    median_size = median(SIZE),
    sd_size = sd(SIZE),
    sample_size = length(SIZE)
  )
abund_napl_table
# way more 2017 observations than 2012, is that a problem?
```

```{r size qqplots, include = FALSE}
size_qq_ivee <- ggplot(abund_ivee, aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ YEAR)
size_qq_ivee
# kinda looks normal?

size_qq_napl <- ggplot(abund_napl, aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ YEAR)
size_qq_napl
# kinda looks normal?
```

```{r size histograms, include = FALSE}
size_hist_ivee <- ggplot(abund_ivee, aes(x = SIZE)) +
  geom_histogram(bins = 17)
size_hist_ivee
# not normally distributed

size_hist_napl <- ggplot(abund_napl, aes(x = SIZE)) +
  geom_histogram(bins = 10)
size_hist_napl
# normally distributed
```

```{r test for equal variances, include = FALSE}
size_ivee_2012 <- abund %>%
  filter(SITE == "IVEE",
         YEAR == "2012") %>% 
  pull(SIZE)

size_ivee_2017 <- abund %>% 
  filter(SITE == "IVEE",
         YEAR == "2012") %>% 
  pull(SIZE)

size_napl_2012 <- abund %>%
  filter(SITE == "NAPL",
         YEAR == "2012") %>% 
  pull(SIZE)

size_napl_2017 <- abund %>% 
  filter(SITE == "NAPL",
         YEAR == "2017") %>% 
  pull(SIZE)

ivee_var <- var.test(size_ivee_2012, size_ivee_2017)
ivee_var
# what does p-value of 1 mean???
napl_var <- var.test(size_napl_2012, size_napl_2017)
napl_var
# p value 0.7685, variances are equal so use Student's not Welch
```

```{r t-test, include = FALSE}
# idk what's going on with IVEE but did it anyway
ivee_t <- t.test(size_ivee_2012, size_ivee_2017, var.equal = TRUE)
ivee_t
# p value = 1?? lob means are not significantly different between 2012 and 2017 at IVEE

napl_t <- t.test(size_napl_2012, size_napl_2017, var.equal = TRUE)
napl_t
# p value = 0.5373, lob means are not significantly different between 2012 and 2017 at NAPL

# In a report, maybe: Between 2012 and 2017, mean lobster size did not change significantly at either Isla Vista (t(df) = statistic, p = statistic, a = 0.05) or Naples Reef (t(df) = statistic, p = statistic, a = 0.05).
```

##### stats n shit: non-MPA sites (part 3)
```{r nonMPA summary tables, include = FALSE}
# Are mean lobster sizes significantly different in 2012 and 2017 for AQUE? MOHK? CARP?
# H0: means are not significantly different in 2012 and 2017 for AQUE/MOHK/CARP.
# HA: means are significantly different in 2012 and 2017 for AQUE/MOHK/CARP.

# 1. establish normality: is 2012 IVEE/2017 IVEE and 2012 NAPL/2017 NAPL data normal?
# 2. test for equal variances for Student's or Welch
# 3. run a t-test

abund_aque <- abund %>% 
  filter(YEAR == "2012" | YEAR == "2017",
         SITE == "AQUE")

abund_mohk <- abund %>% 
  filter(YEAR == "2012" | YEAR == "2017",
         SITE == "MOHK")

abund_carp <- abund %>% 
  filter(YEAR == "2012" | YEAR == "2017",
         SITE == "CARP")

abund_aque_table <- abund_aque %>% 
  group_by(YEAR) %>% 
  summarize(
    mean_size = round(mean(SIZE),2),
    median_size = round(median(SIZE),2),
    sd_size = round(sd(SIZE),2),
    sample_size = length(SIZE)
  )
abund_aque_table

abund_mohk_table <- abund_mohk %>% 
  group_by(YEAR) %>% 
  summarize(
    mean_size = round(mean(SIZE),2),
    median_size = round(median(SIZE),2),
    sd_size = round(sd(SIZE),2),
    sample_size = length(SIZE)
  )
abund_mohk_table

abund_carp_table <- abund_carp %>% 
  group_by(YEAR) %>% 
  summarize(
    mean_size = round(mean(SIZE),2),
    median_size = round(median(SIZE),2),
    sd_size = round(sd(SIZE),2),
    sample_size = length(SIZE)
  )
abund_carp_table
```

```{r nonMPA qqplots, include = FALSE}
size_qq_aque <- ggplot(abund_aque, aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ YEAR)

size_qq_aque
# looks normal

size_qq_mohk <- ggplot(abund_mohk, aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ YEAR)

size_qq_mohk
# looks normal

size_qq_carp <- ggplot(abund_carp, aes(sample = SIZE)) +
  geom_qq() +
  facet_wrap(~ YEAR)

size_qq_carp
# looks normal
```

```{r nonMPA histograms, include = FALSE}
size_hist_aque <- ggplot(abund_aque, aes(x = SIZE)) +
  geom_histogram(bins = 9)

size_hist_aque
# normal ish?

size_hist_mohk <- ggplot(abund_mohk, aes(x = SIZE)) +
  geom_histogram(bins = 13)

size_hist_mohk
# looks normal

size_hist_carp <- ggplot(abund_carp, aes(x = SIZE)) +
  geom_histogram(bins = 13)

size_hist_carp
```

##### lob mean comparisons (part 2)
```{r lob size ANOVA, include = FALSE}
size_anova <- aov(SIZE ~ SITE, data = abund_2017)
summary(size_anova)

size_ph <- TukeyHSD(size_anova)
size_ph
```

```{r test for variances of lob size, include = FALSE}
CL_variances <- abund_2017 %>% 
  group_by(SITE) %>% 
  summarize(
    lob_var = var(SIZE)
  )

CL_variances
```

