---
title: "Final report"
author: "An Bui, Sam Csik"
date: "11/14/2018"
output:
  word_document: default
editor_options:
  chunk_output_type: console
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Introduction

  The California spiny lobster, *Panulirus interruptus*, ranges from Monterey Bay, California, to the southern tip of the Baja California peninsula in temperate-subtropical waters (Lindberg 1955, Vega Velázquez 2003). As adults, they inhabit rocky reef, intertidal, and kelp forest habitats, emerging from shelters at night to forage primarily on benthic invertebrates including mussels (Robles et al. 1990) and red and purple urchins (Tegner & Levin 1983). Within southern California nearshore ecosystems, lobsters can have important impacts as keystone predators. They have been shown to increase biodiversity within rocky intertidal regions in Southern California by limiting the density and sizes of mussels (Robles 1987, Robles et al. 1990, Robles & Robb 1993, Robles 1997, Robles et al. 2001). They may also indirectly facilitate the resistance of giant kelp forests from phase shifts to urchin barrens by consuming red and purple urchins (Tegner & Levin 1983, Halpern et al. 2006, Eurich et al. 2014).  
  
  Beyond their ecological importance, the CA spiny lobster supports a valuable commercial fishery extending from Point Conception to the US-Mexico border, generating over $18 million in ex-vessel landings annually (CDFW 2016). An avid recreational fishery also exists, extending further north to San Luis Obispo County and contributes an estimated 40 million in cosumer spending to the state's economy each year (CDFW 2016). Restrictions on commercial and recreational takes include (1) a minimum carapace length (CL) limit (82.6 mm), which allows most individuals to reproduce for up to two years before reaching the legal size limit and (2) a seasonal fishery closure (March - Ocotober), which protects females from harvest during the species' spawning period. These management measures have helped to promote a stable lobster stock since 2000, though an increase in commercial fishing effort raises questions about the long-term sustainability of the fishery.
  
  In order to maintain healhty lobster populations in the Santa Barbara Channel, as well as the kelp forest communities that they support, studies have recommended the establishment of no-take marine reserves to supplement traditional fisheries management practices (Airame et al. 2003). The establishment of such protected sites in 2012 has provided the Santa Barbara Coastal Long Term Ecological Reserach Project a unique opportunity to monitor lobster populations under different protection regimes and  investigate the effects of fishing on kelp forest community dynamics (Reed 2017). Data collected by the SBC LTER can be used by researchers and policy makers to assess both the impacts of fishing efforts on coastal kelp forest communitites, as well as the effectiveness of marine protected areas (MPAs) in promoting healthy lobster populations. Studies have shown that only larger lobsters (>90 mm) are capable of consuming adult urchins (Tegner & Dayton 1981). Because areas with predominantly adult urchin assemblages are more prone to becoming urchin barrens (Mann 1977, Brietburg 1996), the protection of lobsters above the legal size limit may be imperative in sustaining the health of giant kelp. As such, it is important that researchers and govenrment agencies are knowledgeable of the size structure and abundance of these important predators within coastal kelp forest habitats if they are to effectively manage their populations, habitat, and fisheries.

#### Methods and objectives
##### Marine Protected Areas in the SBC LTER
  California's Marine Life Protection Act was passed in 1999 to designate a network of marine protected areas with the input of scientists and citizen stakeholders. An express purpose of the act was to practice "adaptive management", or management of ecosystems and their communities by integrating current science to continually change management strategies to best fit the needs of the system as well as the stakeholders (cite MLPA).  
  In the five focal sites within the Santa Barbara Coastal Long Term Ecological Research (SBC LTER) sites, two of these are MPAs (Isla Vista, Naples Reef), while three (Arroyo Quemado, Carpinteria, Mohawk) are not. Isla Vista and Naples Reef were designated MPAs in 2012. This design is a natural experiment to understand the efficacy of fishing pressure on the population dynamics of integral species to the kelp forest, such as the CA spiny lobster. When integrating lobster surveys with algal and invertebrate surveys also being conducted at these sites, we are able to construct a holistic profile of kelp forest community dynamics in response to fishing pressure. Additionally, since surveys began in the same year that Isla Vista and Naples Reef were designated MPAs, we are able to assess the length of time before seeing changes in population dynamics in response to eliminating fishing pressure.

##### Long term surveys
  Lobster surveys began in 2012 and were conducted annually as part of a long-term monitoring effort within Santa Barbara Coastal Long Term Ecological Research (SBC LTER) sites. Lobster count and sizes are recorded by divers on SCUBA on four 60 x 5m transects approximately 1km offshore at five sites in the SBC LTER (Fig. 1).  
  Fishing pressure is quantified by counting floating trap buoys at eight SBC LTER sites every two weeks during the legal lobster fishing season from October to March. Trap buoys are attached to wire lobster traps set no deeper than 15m below the ocean surface.

##### Objectives
The following report analyzes lobster abundances and sizes at five SBC LTER locations, two of which are MPAs (Isla Vista, Naples Reef), and three non-MPAs (Arroyo Quemado, Carpenteria, Mohawk). We use monitoring data at these five sites to ask:
    1. how lobster abundance, size, and fishing pressure has changed over time between the start of surveys in 2012 and the most recent surveys in 2017,  
    2. what proportion of lobsters are over the legal size minimum of 82.6mm for the commercial fishery in 2017, and  
    3. how lobster mean size compares across all five sites in 2017.  
We find that _____________. 

#### Results and Discussion

- importance of size/what means for reproduction; prelim data from wet lab (Stier et al., unpublished data) suggests that only medium - large lobsters will consume very small urchins (contrary to everyone wanting to believe that lobsters are urchin-eating machines)...fishing out legal/larger-than-legal lobsters has implications for their ability to effect urhcin populations-- I THINK I MIGHT SAVE THIS FOR THE DISCUSSION


```{r set up packages and data frames, include = FALSE}
# load required packages
library(Rmisc)
library(vcdExtra)
library(tidyverse)
library(ggsignif)
library(effsize)
library(car)
library(kableExtra)
library(onewaytests)
library(reshape2)
library(lemon)
library(ggpubr)

# read in lobster_size_abundance.csv file and get into tidy format
size_abund <- read_csv("lobster_size_abundance.csv")

abund <- as.data.frame(size_abund) %>% 
  expand.dft(freq = "COUNT") # expands freq. table to df representing individual obs. in the table

# read in lobster_traps.csv file and get into tidy format
traps_master <- read_csv("lobster_traps.csv")

traps <- as.data.frame(traps_master) %>% 
  expand.dft(freq = "TRAPS")
```

##### Lobster abundance and fishing pressure between 2012 and 2017

```{r trends in lob abundance and fishing pressure, echo = FALSE}
# PART 1. DESCRIBE TRENDS IN LOBSTER ABUNDANCE AND FISHING PRESSURE AT THE 5 LOCATIONS FROM 2012-2017

#############################
#  calculate mean abundance by year for each site
##############################
abund_by_yearssite <- abund %>% 
  group_by(YEAR, SITE) %>% # group by year, then site
  summarize(
    count = length(SIZE) # sample size by site
  ) %>% 
  mutate(site_full = case_when( # expand to full length names
      SITE == "AQUE" ~ "Arroyo Quemado",
      SITE == "CARP" ~ "Carpinteria",
      SITE == "IVEE" ~ "Isla Vista",
      SITE == "MOHK" ~ "Mohawk Reef",
      SITE == "NAPL" ~ "Naples Reef"
    )
  ) %>% 
  select(YEAR, SITE, site_full, count) # select and reorder columns

##############################
#  calculate mean fishing pressure by year for each site
##############################

traps_by_yearssite <- traps %>% 
  select(YEAR, SITE, SWATH_START, SWATH_END) %>% 
  filter(SITE == "AQUE" | SITE == "CARP" | SITE == "IVEE" | SITE == "MOHK" | SITE == "NAPL") %>% 
  group_by(YEAR, SITE) %>% # group by year, then site
  summarize(
    count = length(SWATH_START) # number of traps by site
  ) %>% 
  mutate(site_full = case_when( # expand to full length names
      SITE == "AQUE" ~ "Arroyo Quemado",
      SITE == "CARP" ~ "Carpinteria",
      SITE == "IVEE" ~ "Isla Vista",
      SITE == "MOHK" ~ "Mohawk Reef",
      SITE == "NAPL" ~ "Naples Reef"
    )
  ) %>% 
  select(YEAR, SITE, site_full, count) # select and reorder columns

##############################
# plot trends in lobster abundance AND fishing pressure by site thorugh time TOGETHER on same plot
##############################

abund_by_yearssite$site_full <- factor(abund_by_yearssite$site_full, levels=c("Isla Vista", "Naples Reef","Arroyo Quemado", "Carpinteria", "Mohawk Reef"))

traps_by_yearssite$site_full <- factor(traps_by_yearssite$site_full, levels=c("Isla Vista", "Naples Reef","Arroyo Quemado", "Carpinteria", "Mohawk Reef"))

caption1 <- paste(strwrap("Figure 1. Lobster abundance and fishing pressure at five locations in the Santa Barbara Channel (2017). Lobster abundance (solid lines) estimated annually on transects conducted via SCUBA at five locations. Fishing pressure is estimated by counting the number of lobster trap floats (dashed lines) in definded ares of coastline every 2 - 4 weeks during the fishing season. MPAs (green) were designated in 2012. Unprotected areas (yellow) are fished annually. Data from Reed (2017).", width = 110), collapse = "\n")

#pdf("fig_1.pdf", width = 8, height = 6)
abund_traps_plot <- ggplot(abund_by_yearssite, aes(x = YEAR, y = count)) +
  geom_point(color = "gray30") +
  geom_line(color = "gray30") +
  geom_point(data = traps_by_yearssite, color = "gray30") +
  geom_line(data = traps_by_yearssite, linetype = "dashed", color = "gray30") +
  theme_classic() +
  theme(axis.text = element_text(color = "black"), 
        panel.border = element_rect(colour = "black", fill=NA, size=0.7), 
        strip.background = element_rect(fill = "grey"),
        plot.caption = element_text(size = 10, hjust = 0)) +
  facet_grid(~site_full) +
  labs(x = "Year", y = "Counts", caption = caption1) +
  scale_x_continuous(breaks=c(2013, 2015, 2017))
abund_traps_plot
#dev.off()
```

##### Lobster size in 2017 inside and outside MPAs
```{r compare mean lob size in 2017}
# PART 2. COMPARE MEAN LOBSTER SIZE BY SITE IN 2017

##############################
# wrangle data for 2017
##############################

abund_2017 <- abund %>% 
  filter(YEAR == "2017",
         SIZE != "-99999") %>% 
  mutate(site_full = case_when( # expand to full length names
      SITE == "AQUE" ~ "Arroyo Quemado",
      SITE == "CARP" ~ "Carpinteria",
      SITE == "IVEE" ~ "Isla Vista",
      SITE == "MOHK" ~ "Mohawk Reef",
      SITE == "NAPL" ~ "Naples Reef"
    )
  ) %>% 
  select(YEAR, SITE, site_full, SIZE)

##############################
# test for variances, ANOVA, Tukey's HSD
##############################

lob_variances <- abund_2017 %>% 
  group_by(SITE) %>% 
  summarize(
    lob_var = var(SIZE)
  )

lob_variances 
# largest variance is no greater than 4x smallest variance

size_anova <- aov(SIZE ~ SITE, data = abund_2017)
summary(size_anova)

size_ph <- TukeyHSD(size_anova)
size_ph  # so the only ones that are different are NAPL-CARP and NAPL-IVE

##############################
# box and whisker plot of size
##############################

abund_2017$site_full <- factor(abund_2017$site_full, levels=c("Isla Vista", "Naples Reef","Arroyo Quemado", "Carpinteria", "Mohawk Reef")) # make site_full a factor to reorder on plot

caption2 <- paste(strwrap("Figure 2. Carapace lengths (CL, mm) of CA spiny lobster at five locations in the Santa Barbara Channel (2017). CL estimated on SCUBA in study transects at five locations: Isla Vista, Naples Reef, Arroyo Quemado, Carpenteria, Mohawk. MPAs (green) were designated in 2012. Unprotected areas (yellow) are fished annually. Dashed line represents the mean CL across all five sites in 2017. Mean CL only differed significantly between Naples Reef-Isla Vista (_p_ = 0.004) and Naples Reef-Carpenteria (_p_ = 0.02). All other pairwise comparisons yielded no signficant differences (one-way ANOVA, F(4) = 3.42, p < 0.01) with post-hoc Tukey's HSD (alpha = 0.05). Data from Reed (2017).", width = 110), collapse = "\n")

#stats <- summarySE(abund_2017, measurevar = "SIZE", groupvars = c("site_full"), na.rm = TRUE) # also need to insert "stats" into ggplot argument, but isn't working correctly

pdf("fig_2.pdf", width = 8, height = 6)
size_2017_boxplot <- ggplot(abund_2017, aes(x = site_full, y = SIZE)) +
  geom_boxplot(fill = "darkgrey") +
  geom_hline(yintercept = 82.6, linetype = 2) + # add line at CL mean across sites
  geom_signif(comparisons = list(c("Naples Reef", "Isla Vista"), c("Naples Reef", "Carpinteria")),
              map_signif_level = TRUE) +
  theme_classic() +
  theme(axis.text = element_text(color = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=0.8),
        legend.position = "none",
        plot.caption = element_text(size = 10, hjust = 0)) +
  labs(x = "Site", y = "Carapace Length (mm)", caption = caption2)

size_2017_boxplot
dev.off()
```

##### Lobster size change over time at MPA sites
```{r comparing lob sizes in 2012 and 2017}
# PART 3. AT IVEE AND NAPL (MPAs), HOW DO LOBSTER SIZES IN 2012 AND 2017 COMPARE? AT NON-MPA SITES?

##############################
# subset data by site and year (2012, 2017) to compare sizes at IVEE and NAPL
##############################

size_ivee_2012 <- abund %>%
  filter(SITE == "IVEE",
         YEAR == "2012") %>% 
  pull(SIZE)

size_ivee_2017 <- abund %>% 
  filter(SITE == "IVEE",
         YEAR == "2017") %>% 
  pull(SIZE)

size_napl_2012 <- abund %>%
  filter(SITE == "NAPL",
         YEAR == "2012") %>% 
  pull(SIZE)

size_napl_2017 <- abund %>% 
  filter(SITE == "NAPL",
         YEAR == "2017") %>% 
  pull(SIZE)

##############################
# one-way t-test for IVEE & NAPL
##############################

ivee_t <- t.test(size_ivee_2012, size_ivee_2017, var.equal = TRUE) 
ivee_t # p value = 0.0599, lob means are not sig different between 2012 & 2017 at IVEE (but close to sig...)

napl_t <- t.test(size_napl_2012, size_napl_2017, var.equal = TRUE)
napl_t # p value = 0.5002, lob means are not significantly different between 2012 and 2017 at NAPL
```

  Between 2012 and 2017, mean lobster size does not differ signficantly at either Isla Vista (t(`r round(ivee_t$parameter, 2)`) = `r round(ivee_t$statistic, 2)`, _p_ = `r round(ivee_t$p.value, 2)`) or Naples Reef (t(`r round(napl_t$parameter, 2)`) = `r round(napl_t$statistic, 2)`, _p_ = `r round(napl_t$p.value, 2)`). 
  
```{r lob sizes at non-MPA sites}
##############################
# subset data by site and year (2012, 2017) to compare sizes at non-MPA sites (AQUE, CARP, MOHK)
##############################

size_aque_2012 <- abund %>%
  filter(SITE == "AQUE",
         YEAR == "2012") %>% 
  pull(SIZE)

size_aque_2017 <- abund %>% 
  filter(SITE == "AQUE",
         YEAR == "2017") %>% 
  pull(SIZE)

size_carp_2012 <- abund %>%
  filter(SITE == "CARP",
         YEAR == "2012") %>% 
  pull(SIZE)

size_carp_2017 <- abund %>% 
  filter(SITE == "CARP",
         YEAR == "2017") %>% 
  pull(SIZE)

size_mohk_2012 <- abund %>%
  filter(SITE == "MOHK",
         YEAR == "2012") %>% 
  pull(SIZE)

size_mohk_2017 <- abund %>% 
  filter(SITE == "MOHK",
         YEAR == "2017") %>% 
  pull(SIZE)

##############################
# two-way t-test for non-MPA sites
##############################

aque_t <- t.test(size_aque_2012, size_aque_2017, var.equal = TRUE) # p = 0.2097
carp_t <- t.test(size_carp_2012, size_carp_2017, var.equal = TRUE) # p = 0.1819
mohk_t <- t.test(size_mohk_2012, size_mohk_2017, var.equal = TRUE) # p < 0.001

# Similarly to MPA sites, non-MPA sites show a similar consistency in lobster carapace length when comparing measurements from 2012 and 2017. Carapace length does not differ significantly between 2012 and 2017 at Arroyo Quemado (t(`r round(aque_t$parameter, 2)`) = `r round(aque_t$statistic, 2)`, _p_ = `r round(aque_t$p.value, 2)`) or Carpinteria (t(`r round(carp_t$parameter, 2)`) = `r round(carp_t$statistic, 2)`, _p_ = `r round(carp_t$p.value, 2)`). However, lobster carapace length does differ significantly between comparison years at Mohawk (t(`r round(mohk_t$parameter, 2)`) = `r round(mohk_t$statistic, 2)`, _p_ = `r round(mohk_t$p.value, 2)`). 

##############################
# wrangle data to make boxplot comparing 2012 & 2017 for each site
##############################

abund_df <- abund %>% 
  group_by(SITE, YEAR) %>% # group by year, then site
  filter(YEAR == "2012" | YEAR == "2017") %>% 
  mutate(protection = case_when(
    SITE == "NAPL" ~ "MPA",
    SITE == "IVEE" ~ "MPA",
    SITE == "AQUE" ~ "Non-MPA",
    SITE == "CARP" ~ "Non-MPA",
    SITE == "MOHK" ~ "Non-MPA")) %>% 
  mutate(site_full = case_when( 
      SITE == "AQUE" ~ "Arroyo Quemado",
      SITE == "CARP" ~ "Carpinteria",
      SITE == "IVEE" ~ "Isla Vista",
      SITE == "MOHK" ~ "Mohawk Reef",
      SITE == "NAPL" ~ "Naples Reef"
    )) %>% 
  select(YEAR, SITE, site_full, SIZE, protection)

##############################
# boxplot of sizes in 2012 & 2017 in MPA & non-MPA sites
##############################

abund_df$site_full <- factor(abund_df$site_full, levels=c("Isla Vista", "Naples Reef","Arroyo Quemado", "Carpinteria", "Mohawk Reef"))

caption3 <- paste(strwrap("Figure 3. Carapace lengths (CL, mm) of CA spiny lobster at five locations in the Santa Barbara Channel (2012 & 2017). Mean carapace length differed significantly between 2012 and 2017 at Mohawk Reef (t(259) = 4.07, p < 0.001) only. Data from Reed (2017).", width = 110), collapse = "\n") # fix alpha; bold; how to discuss statistical results?

pdf("fig_3.pdf", width = 8, height = 6)
abund_plot <- ggplot(abund_df, aes(x = YEAR, y = SIZE)) +
  geom_boxplot(aes(group = YEAR, fill = factor(YEAR))) +
  scale_fill_manual(breaks = c("2012", "2017"),values = c("white", "darkgrey"), name = "Year") +
  facet_grid(~ site_full) +
  labs(x = "Site", y = "Carapace Length (mm)", caption = caption3) +
  theme_classic() +
   theme(
    axis.title.x=element_blank(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=0.7),
    strip.background = element_rect(fill = "grey"),
    plot.caption = element_text(size = 10, hjust = 0)) 
abund_plot
dev.off()

##############################
# Cohen's d effect size -- change in CL from 2012-2017 at each site
##############################

napl_d <- cohen.d(size_napl_2012, size_napl_2017, pooled = TRUE) # SMALL increase
ivee_d <- cohen.d(size_ivee_2012, size_ivee_2017, pooled = TRUE) # SMALL increase
aque_d <- cohen.d(size_aque_2012, size_aque_2017, pooled = TRUE) # SMALL increase
carp_d <- cohen.d(size_carp_2012, size_carp_2017, pooled = TRUE) # NEGLIGIBLE decrease
mohk_d <- cohen.d(size_mohk_2012, size_mohk_2017, pooled = TRUE) # MEDIUM decrease

##############################
# absolute differences
##############################

mean_ivee_2012 <- mean(size_ivee_2012)
mean_ivee_2017 <- mean(size_ivee_2017)
change_ivee <- mean_ivee_2017 - mean_ivee_2012 # 5.375 mm increase

mean_napl_2012 <- mean(size_napl_2012)
mean_napl_2017 <- mean(size_napl_2017)
change_napl <- mean_napl_2017 - mean_napl_2012 # 3.232 mm increase

mean_aque_2012 <- mean(size_aque_2012)
mean_aque_2017 <- mean(size_aque_2017)
change_aque <- mean_aque_2017 - mean_aque_2012 # 2.8955 mm increase

mean_carp_2012 <- mean(size_carp_2012)
mean_carp_2017 <- mean(size_carp_2017)
change_carp <- mean_carp_2017 - mean_carp_2012 # 2.129 mm decrease

mean_mohk_2012 <- mean(size_mohk_2012)
mean_mohk_2017 <- mean(size_mohk_2017)
change_mohk <- mean_mohk_2017 - mean_mohk_2012 # 5.253 mm decrease
```

```{r proportion of lobs above legal min}
# PART 4. WHAT PROPORTION OF OBSERVED LOBSTERS AT EACH SITE ARE ABOVE THE LEGAL MINIMUM (82.6mm)

##############################
# make new table with sizes above and below legal limit
##############################

abund_2017_size <- abund_2017 %>% 
  mutate(legal = case_when(
    SIZE > 82.6 ~ "above",
    SIZE < 82.6 ~ "below"
  )) %>% 
  count(SITE, legal) %>% 
  spread(legal, n) %>% 
  select(-SITE)

rownames(abund_2017_size) <- c("Arroyo Quemado", "Carpinteria", "Isla Vista Reef", "Mohawk", "Naples Reef")
  
##############################
# x2 test
##############################

lob_prop <- round(prop.table(as.matrix(abund_2017_size), 1), 3)*100
lob_prop

lob_prop_columns = c("Above", "Below")

lob_prop_caption = "**Table 2. Proportions of lobsters above and below legal limit in 2017.** The legal length of fished lobsters is 82.6mm."

kable(lob_prop,
      col.names = lob_prop_columns,
      caption = lob_prop_caption) %>% 
  kable_styling(bootstrap_options = "striped")

# how to add column name to site? need to create new data frame?
# going to try to figure out how to add a table that has counts with proportions

lob_x2 <- chisq.test(abund_2017_size)
lob_x2
# The proportion of legal size lobsters is significantly different amongst sites ($\chi^2$(`r lob_x2$parameter)) = `r round(lob_x2$statistic, 2`), _p_ < 0.001). 
```

###References

1. Lindberg RJ (1955) Growth, population dynamics, and field behavior in the spiny lobster (Panulirus interruptus) (Randall). University of California Publications in Zoology. 59:157-248.
2. Velazquez (with accent over "a") AV (2003) Reproductive strategies of the spiny lobster (Panulirus interruptus) related to the marine environmental variability off central Baja California, Mexico: management implications. Fishieries Research. 65:123-135.
3. Robles C, Sweetnam D, Eminike J (1990) Lobster predation on mussels: shore-level differences in prey vulnerability and predator preference. Ecology 71:1564-1577.
4. Tegner MJ & Levin LA (1983) Spiny lobsters and sea urchins: analysis of a predator-prey interaction. Journal of Experimental Marine Biology and Ecology. 73:125-150.
5. Robles C (1987) Predator foraging characteristics and prey population structure on a sheltered shore. Ecology. 68:1502-1514
6. Robles C & Robb J (1993) Varied carnivore effects and the prevalence of intertidal algal turfs. Journal of Experimental Marine Biology and Ecology. 166:65-91.
7. Robles C (1997) Changing recruitment in constant species assemblages: implications for predation theory in intertidal communities. Ecology. 78:1400-1414
8. Robles CD, Alvarado MA, Desharnais RA (2001) The shifting balance of littoral predator-prey interaction in regimes of. hydrodynamic stress. Oecologia. 128:142-152
9. Halpern BS, Cottenie K, Broitman BR (2006) Strong top-down control in southern California kelp forest ecosystems. Science. 312:1230-1232.
10. Eurich JG, Seldon RL, Warner RR (2014) California spiny lobster preference for urchins from kelp forests: implications for urchin barren persistence. Marine Ecology Progress Series. 498:217-225.
11. California Department of Fish & Wildlife (2016) California spiny lobster fishery managment plan.
12. Airame S, Dugan JE, Lafferty KD, Leslie H, McArdle DA, Warner RR (2003) Applying ecological criteria to marine reserve design: a case study from the California Channel Islands. Ecological Applications. 13:S170-S184.
Airman et al. 2003
13. Reed D (2017) SBC LTER: Reef: Abundance, size and fishing effort for California spiny lobster (Panulirus interruptus), ongoing since 2012. Santa Barbara Coastal Long Term Ecological Research Project. doi:10.6073/pasta/81ce20b29614ec99d85d54907eaa3e8e
14. Tegner MJ & Dayton PK (1981) Population structure, recruitment and mortality of two sea urchins (Strogylocentrotus franciscanus an S. purpuratus) in a kelp forest. Marine Ecology Progress Series 5:255-268.
15. Mann KH (1977) Destruction of kelp-beds by sea urchins: a cyclical phenomenon or irreversible degradation? Helgoländer wiss. Merresunters. 30:455-467.
16. Breitburg DL (1996) Consumer mobility and the relative importance of consumption and competition following physical disturbance. Marine Ecology Progress Series. 138:83-92.

